\chapter{Implementation Details}
\label{ch:implementation}

This section translates the algorithms of Chapter~\ref{ch:methodology} into concrete artefacts: the proxy server, the typed TypeScript SDK, the backend service, and the Cloud Run deployment.

\section{Repository Layout}
\label{sec:repo}

The platform is organised as a hybrid monorepo:
\begin{itemize}
    \item \texttt{sprout-web/} --- Next.js 14 frontend (TypeScript, Tailwind, Shadcn/ui, Zustand). Static export deployed to Firebase Hosting.
    \item \texttt{sprout-backend/} --- Firebase Cloud Functions (Node.js + TypeScript). REST API, Firestore triggers, auth middlewares.
    \item \texttt{gemini-proxy/} --- Stateless Node.js WebSocket gateway on Cloud Run. No per-user data beyond the in-process session map.
    \item \texttt{packages/gemini-proxy-sdk/} --- Published as \texttt{@abyxolv/gemini-proxy-sdk}.
    \item \texttt{sprout-app/} --- Terraform provisioning Firebase, Cloud Run, Cloud Tasks, and secrets~\cite{firebase2024,cloudrun2024}.
\end{itemize}

\section{Gemini Proxy Server}
\label{sec:proxy}

The proxy is built around the \texttt{ws} WebSocket library. Listing~\ref{lst:jwt} shows JWT validation; Listing~\ref{lst:session} the session manager; Listing~\ref{lst:backend} the backend event client.

\begin{lstlisting}[caption={JWT validation middleware (\texttt{auth-middleware.js}).},label=lst:jwt]
const { OAuth2Client } = require('google-auth-library');
const client = new OAuth2Client(process.env.GOOGLE_CLIENT_ID);

async function validateGoogleJWT(token) {
  const ticket = await client.verifyIdToken({
    idToken: token,
    audience: process.env.GOOGLE_CLIENT_ID,
  });
  return ticket.getPayload();
}
\end{lstlisting}

\begin{lstlisting}[caption={Session manager (\texttt{session-manager.js}).},label=lst:session]
class SessionManager {
  constructor() {
    this.sessions = new Map();
    this.userSessions = new Map();
  }
  createSession(uid, userData) {
    const sessionId = generateSessionId();
    const session = {
      id: sessionId, uid, userData,
      startTime: Date.now(),
      tokenUsage: { input: 0, output: 0 },
      messageCount: 0, status: 'active'
    };
    this.sessions.set(sessionId, session);
    this.userSessions.set(uid, sessionId);
    return session;
  }
}
\end{lstlisting}

\begin{lstlisting}[caption={Backend event client.},label=lst:backend]
class BackendClient {
  async sendSessionEvent(sessionId, eventType, data) {
    try {
      await axios.post(
        `${this.baseUrl}/api/sessions/${sessionId}/events`,
        { type: eventType, data, timestamp: Date.now() }
      );
    } catch (e) {
      console.error('Failed to send backend event:', e);
    }
  }
}
\end{lstlisting}

\section{Client SDK}
\label{sec:sdk}

The frontend uses \texttt{@abyxolv/gemini-proxy-sdk}, which standardises authentication, reconnection, audio framing, and metric capture. Listing~\ref{lst:sdk-api} shows the public TypeScript surface; Listing~\ref{lst:browser} typical browser usage.

\begin{lstlisting}[caption={Public SDK API.},label=lst:sdk-api]
export interface TokenProvider {
  getToken(): Promise<string>;
}
export interface GeminiProxyClientOptions {
  proxyUrl: string;
  tokenProvider: TokenProvider;
  reconnect?: { enabled: boolean; maxRetries?: number };
}
export interface GeminiProxyClient {
  connect(): Promise<void>;
  disconnect(code?: number, reason?: string): void;
  send(data: unknown): void;
  sendText(text: string): void;
  on(listener: (e: GeminiEvent) => void): () => void;
}
\end{lstlisting}

\begin{lstlisting}[caption={Browser usage.},label=lst:browser]
const client = createGeminiClient({
  proxyUrl: process.env.NEXT_PUBLIC_GEMINI_PROXY_URL!,
  tokenProvider: { getToken: () => auth.currentUser!.getIdToken() },
  reconnect: { enabled: true, maxRetries: 5, backoffMs: 1000 },
});
await client.connect();
client.on((e) => {
  if (e.type === 'message') { /* update UI */ }
  if (e.type === 'usage') { /* update token meter */ }
});
\end{lstlisting}

\section{Backend Service Surface}
\label{sec:backend-api}

Public endpoints include \texttt{POST /onboarding/skills}, \texttt{POST /courses/knowledge-map}, \texttt{GET /courses/\{id\}/lessons/\{id\}/completion}, and \texttt{POST /conversations/feedback}. Internal proxy endpoints include \texttt{POST /internal/session/create-session} and \texttt{POST /internal/chat/events/callback}. The split ensures the live data plane never invokes user-scoped auth paths.

\section{Cloud Run Deployment}
\label{sec:cloudrun}

Listing~\ref{lst:cloudrun} shows the service definition: container concurrency 50, CPU throttling disabled, minimum instances 1 to avoid cold-start latency.

\begin{lstlisting}[caption={Cloud Run service definition for \texttt{gemini-proxy}.},label=lst:cloudrun]
apiVersion: serving.knative.dev/v1
kind: Service
metadata:
  name: gemini-proxy
spec:
  template:
    metadata:
      annotations:
        run.googleapis.com/cpu-throttling: 'false'
        run.googleapis.com/min-instances: '1'
        run.googleapis.com/max-instances: '200'
    spec:
      containerConcurrency: 50
      containers:
      - image: gcr.io/$PROJECT_ID/gemini-proxy:latest
        ports:
        - name: h2c
          containerPort: 8080
\end{lstlisting}

\section{Configuration and Secrets}
\label{sec:secrets}

Credentials are sourced from Google Secret Manager. Non-sensitive configuration flows from Terraform variables; \texttt{NEXT\_PUBLIC\_*} variables are injected at build time and are not secrets.

\section{Frontend Components}
\label{sec:frontend}

The AI surface is concentrated in \texttt{AssistantConnector} (SDK lifecycle and modality selection), \texttt{AssistantChat} (transcript and text input), and \texttt{SkillMapChart} (React Flow knowledge map). The \texttt{useLessonCompletionPolling} hook implements Algorithm~\ref{alg:a4} on the Firestore client SDK.
