\chapter{Operations and Reproducibility}
\label{ch:operations}

\section{Environment Separation}
\label{sec:env}

Three environments are maintained: \texttt{dev}, \texttt{uat}, and \texttt{prod}, each in a separate GCP project with isolated Firebase, Firestore, Secret Manager, and Cloud Tasks. Branch mapping: \texttt{develop} $\rightarrow$ dev, \texttt{uat} $\rightarrow$ uat, \texttt{master} $\rightarrow$ prod. GitLab CI runs \texttt{terraform plan/apply} with matching \texttt{<env>.tfvars}.

Table~\ref{tab:env-matrix} summarises how environments differ for the AI subsystem. Production is the only environment that serves anonymous learners at scale; UAT mirrors prod topology with reduced quotas.

\tablesetup
\begin{table}[H]
\centering
\caption{Environment matrix for the AI subsystem.}
\label{tab:env-matrix}
\small
\begin{tabular}{|l|c|c|c|}
\hline
\TblHead{Property} & \TblHead{dev} & \TblHead{uat} & \TblHead{prod} \\
\hline
GCP project & \texttt{sprout-dev} & \texttt{sprout-uat} & \texttt{sprout-prod} \\
Gemini model & Flash (preview) & Flash & Flash / Pro (flag) \\
Proxy \texttt{min-instances} & 0 & 1 & 1 \\
Free-tier minutes / month & 120 & 60 & 60 \\
Mixpanel project & dev & uat & prod \\
Synthetic agent schedule & on deploy & hourly & every 15\,min \\
\hline
\end{tabular}
\end{table}

\section{CI/CD Pipeline}
\label{sec:cicd}

Table~\ref{tab:cicd} lists the GitLab stages that touch the AI subsystem. Proxy and backend deploy independently so a regression in Cloud Functions cannot block a hotfix to WebSocket routing.

\tablesetup
\begin{table}[H]
\centering
\caption{GitLab CI stages for AI-related artefacts.}
\label{tab:cicd}
\small
\begin{tabular}{|p{2.2cm}|p{3.5cm}|p{6.6cm}|}
\hline
\TblHead{Stage} & \TblHead{Trigger} & \TblHead{Actions} \\
\hline
lint & MR to \code{develop} & ESLint on \code{sprout-web}, \code{gemini-proxy}, functions \\
\hline
test & MR to \code{develop} & Jest unit tests; proxy JWT negative suite \\
\hline
build & merge to \code{develop} & Docker image $\rightarrow$ Artifact Registry \\
\hline
terraform & manual approve & \code{plan/apply} per environment \\
\hline
deploy-proxy & post-apply & Cloud Run revision; 10\% canary in prod \\
\hline
deploy-backend & post-apply & Firebase Functions; Firestore rules \\
\hline
deploy-web & post-apply & Static export to Hosting \\
\hline
synthetic & post-deploy & WebSocket smoke test; alert on failure \\
\hline
\end{tabular}
\end{table}

Average lead time from merge to prod proxy deploy is 22 minutes when Terraform plans are clean; rollbacks use traffic shifting to the previous revision within five minutes (Section~\ref{sec:rollback}).

\section{Terraform State}
\label{sec:terraform}

State is held in an HTTP backend on GitLab, one file per environment. Manual GCP console changes are disallowed to prevent drift. Pipelines acquire a state lock for the duration of apply.

\section{Secret Management}
\label{sec:ops-secrets}

Sensitive credentials live in Secret Manager. The proxy mounts \texttt{GOOGLE\_CLIENT\_ID} and \texttt{GEMINI\_API\_KEY} via \texttt{secretKeyRef}. The frontend holds only \texttt{NEXT\_PUBLIC\_*} values.

\section{Health Checks and Synthetic Tests}
\label{sec:health-checks}

The proxy exposes \texttt{/health} with \texttt{status}, \texttt{activeConnections}, \texttt{uptime}, \texttt{memory}, and \texttt{geminiConnection}. Uptime checks poll every 60\,s; a synthetic agent opens a WebSocket, exchanges a known prompt, and asserts a non-empty response.

\section{Alerting}
\label{sec:ops-alerting}

Five Cloud Monitoring rules (Section~\ref{sec:alerting}) page via PagerDuty or Slack with runbook links and recent log queries.

\section{Rollback}
\label{sec:rollback}

Rollback exists at three layers: Cloud Run revision traffic shift; Firebase Functions artefact promotion; Firestore logical schema guarded by feature flags.

\section{Reproducing the Pipeline}
\label{sec:reproduce}

A clean reproduction requires:
\begin{enumerate}
    \item \texttt{terraform apply -var-file=<env>.tfvars} from \texttt{sprout-app}.
    \item Populate Secret Manager (Gemini API key, OAuth client ID, admin keys).
    \item Deploy proxy: \texttt{gcloud builds submit gemini-proxy/} then \texttt{gcloud run deploy gemini-proxy}.
    \item \texttt{firebase deploy --only functions} for backend.
    \item \texttt{npm run build \&\& firebase deploy --only hosting} for frontend.
    \item Run synthetic agent against the deployment.
\end{enumerate}

\section{Determinism Caveats}
\label{sec:determinism}

LLM output is non-deterministic. Cloud Tasks may deliver more than once; handlers are idempotent. WebSocket clocks use the proxy as authority.

\section{Ethics and Compliance}
\label{sec:ethics}

The first tutor interaction displays a recording notice. Minor accounts may disable webcam via admin toggle. Raw audio and video are not persisted. Account deletion triggers cascading delete. Administrative reads are logged for 90 days.
